---
title: "Combine data"
title-block-banner: true
execute:
  message: false
  warning: false
---

```{r}
#| echo: false

library(tidyverse)
library(kableExtra)  # for tables
library(readr)
library(haven)
```

## A single data frame

After having cleaned all individual studies and put the data into a shared format, we combine the studies into a single data frame. 

```{r}
# combine data

# List all matching files recursively
files <- list.files(
  path = "papers", 
  pattern = "^cleaned.*\\.rds$", 
  full.names = TRUE, 
  recursive = TRUE
)

# Load each file into a named list
studies_data <- lapply(files, function(file) {
  readRDS(file) |> 
    # remove old labels from stata files 
    haven::zap_labels() |> 
    # ensure all identifier variables are of same variable type
    mutate(subject_id = as.character(subject_id), 
           scale = as.character(scale))
})

# Name list elements based on filenames (without extension or path)
names(studies_data) <- tools::file_path_sans_ext(basename(files))

data <- bind_rows(studies_data)

# build additional variables
write_csv(data, "data.csv")

```


## Control and intervention group selection

Several experiments have multiple control and/or intervention groups. For our meta-analysis, for each experiment, we only pair one control and one treatment group. In a case where we needed to make a selection, we have coded our selection choice in the variables `intervention_selection` and  `control_selection`. 


```{r}
names(data)


# data |> 
#   filter(
#     # for cases where there was an intervention selection, make sure to only pick the selected intervention
#     is.na(intervention_selection) | intervention_label == intervention_selection,
#     # same but for the control
#     is.na(control_selection) | control_label == control_selection
#   )

```


## Exclude long_term effects

Some studies measured long_term effects. We want to exclude these from the main analysis (including them would likely reduce the detected average treatment effect, as effects tend to regress over time). 






